The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates

### 2024-10-11 _tex/index.tex
Built site for gh-pages

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
sn-nature
]{sn-jnl}

@@

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\raggedbottom
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates},
  pdfauthor={Thomas Gavin Carmichael; Alexander Rauscher; Ruth E Grunau; Alexander Mark Weber},
  pdfkeywords={Quantitative Susceptbility
Mapping, Preterm, Newborn, Cerebral Venous Oxygen Saturation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

% author setup
\author[1,2]{\fnm{Thomas Gavin} \sur{Carmichael}}\email{tgcarmichael@outlook.com}\author[3]{\fnm{Alexander} \sur{Rauscher}}\email{rauscher@physics.ubc.ca}\author[2,3]{\fnm{Ruth E} \sur{Grunau}}\email{rgrunau@mail.ubc.ca}\author*[2,3]{\fnm{Alexander Mark} \sur{Weber}}\email{aweber@bcchr.ca}
% affil setup
\affil[1]{\orgdiv{Integrated Sciences}, \orgname{The University of
British Columbia}, \orgaddress{\street{2329 West
Mall}, \city{Vancouver}, \postcode{V6T 1Z4}, \country{Canada}}}
\affil[3]{\orgdiv{Pediatrics}, \orgname{The University of British
Columbia}, \orgaddress{\street{2329 West
Mall}, \city{Vancouver}, \postcode{V6T 1Z4}, \country{Canada}}}
\affil[2]{\orgdiv{BC Children's Hospital Research
Institute}, \orgname{The University of British
Columbia}, \orgaddress{\street{938 West 28th
Avenue}, \city{Vancouver}, \postcode{V5Z 4H4}, \country{Canada}}}

@@

\abstract{\textbf{Background}: Quantitative susceptibility mapping (QSM)
is a magnetic resonance imaging (MRI) modality proposed to be a viable
method of measuring cerebral oxygenation in neonates given its
sensitivity to deoxyhemoglobin, a paramagnetic molecule. During QSM,
however, paramagnetic sources can be obscured by opposing diamagnetic
sources such as water and myelin. We sought to evaluate whether QSM
images alone, or an algorithm that attempts to isolate their
paramagnetic components, are more accurate in measuring oxygenation of
the major cerebral veins in a cohort of neonates born preterm.
Additionally, we aimed to determine whether a difference in oxygenation
existed between the major cerebral veins.

\textbf{Methods}: 19 neonates born preterm were scanned on a 3T research
MRI at term equivalent age. The protocol included a multi-echo
susceptibility-weighted imaging sequence. The acquired imaging data were
processed as QSM images to obtain the susceptibility values of the
superior sagittal sinus (SSS) and central cerebral veins (CCV). These
values were used to calculate the oxygen saturation
(SvO\textsubscript{2}) of the SSS and CCV. QSM images were subsequently
processed to isolate their paramagnetic components. SvO\textsubscript{2}
values of the SSS and CCV were calculated again from the paramagnetic
components.

\textbf{Results}: The mean SvO\textsubscript{2} values of the SSS and
CCV calculated from QSM images were found to be 72.4\% (SD, 3.4\%) and
68.7\% (SD, 3.5\%), respectively. The mean SvO\textsubscript{2} values
calculated from paramagnetic components were found to be 58.1\% (SD,
7.3\%) for the SSS and 57.7\% (SD, 7.0\%) for the CCV.

\textbf{Conclusion}: SSS SvO\textsubscript{2} values derived from
paramagnetic components agreed well with the existing literature and
were closer than the values derived from QSM, however, they displayed
greater variability. Although the CCV SvO\textsubscript{2} data from QSM
aligns more closely with existing literature, it is important to note
that the current literature on this topic remains relatively limited in
the CCV. Thus, decomposing QSM images into paramagnetic components shows
great promise as a method for more accurately measuring cerebral
oxygenation in neonates but may require more research to improve
precision. Notably, no significant difference in oxygenation was
observed between the CCV and the SSS, contrasting with previous
studies.}

@@

@@

Preterm birth

Abnormal brain development is a significant concern for parents with
children born preterm, as 43\% of infants that survive will have
neurodevelopmental delays later in life (Dix, van Bel, and Lemmers
2017). Irregularities in early cerebral oxygen levels have been
identified as a potential source of such delays, where too little oxygen
provided during NICU care can result in white matter injury, while too
much oxygen can result in reduced cortical connectivity (Rantakari et
al. 2021). As such, being able to precisely, accurately, and
non-invasively measure cerebral oxygenation is necessary for
understanding and improving neurodevelopmental outcomes in preterm
neonates.

In the present study, we set out to determine whether a QSM image alone,
or the paramagnetic component of the QSM image, is more accurate in
measuring the oxygen present in the major cerebral veins of a cohort of
preterm neonates. As previous QSM studies have not included the SSS, we
also had a secondary aim of preserving this vessel in our QSM images and
using this data to determine whether a difference in oxygenation existed
between the SSS and the central cerebral veins (CCV).

@@

Participant data comes from a previous study (Zhu et al. 2024).
Participants consisted of preterm neonates born between 25- and 31-weeks
gestational age (GA) who were admitted to the level III NICU at BC
Women's Hospital. Recruitment took place over a span of one year, from
February 2021 to January 2022, facilitated by a dedicated research
nurse. Parents of eligible infants were approached by the research nurse
prior to discharge from the NICU to explain the study objectives and
seek their consent for participation. Infants meeting the criteria for
inclusion were scanned for the study if they had already been discharged
from the NICU, were in stable condition, and had reached a term
equivalent age of 37 to 44 weeks GA. However, certain exclusion criteria
were applied to ensure the homogeneity and integrity of the study
sample. Infants were excluded if there was clinical evidence of a
congenital malformation or syndrome, a TORCH infection, or ultrasound
evidence of large parenchymal hemorrhagic infarction (\textgreater2 cm,
Grade 4 intraventricular hemorrhage).

@@

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750
scanner (scanner software version DV26.0\_R03) equipped with a SREE
Medical Systems single-channel neonatal head coil (Table~\ref{tbl-mri}).
The scans were conducted at the BC Children's MRI Research Facility.
Prior to the scanning procedure, subjects were carefully prepared by a
research nurse. Swaddling and feeding were used to ensure the comfort
and cooperation of the subjects during the scan. Importantly, no
sedatives or invasive markers were utilized throughout the procedure.
Subjects were placed within a specially designed SREE Medical Systems
MRI compatible incubator, which facilitated both safety and motion
minimization. Molded foam was strategically positioned around the head
and body within the incubator to further restrict subject movement. To
protect against potential hearing damage, ear plugs were employed during
the scanning process. Additionally, a pulse oximeter was affixed to the
subject's foot to monitor arterial oxygen saturation and heart rate
throughout the scan.

@@

\caption{\label{tbl-mri}Technical parameters for MR imaging pulse
sequences}

@@

\toprule
 & T1w & T2w & pcASL & SWI\\
\midrule
Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated\\
Phase-encoding direction & Coronal & Sagittal & Axial & Axial\\
TR (ms) & 7.74 & 2,300 & 4,680 & 30.9\\
TE (ms) & 2.97 & 66.29 & 10.55 & 5.24\\
Flip angle & 12° & 90° & 111° & 20°\\
FOV (cm) & 20 & 20 & 24 & 25\\
Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256\\
In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977\\
Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2)\\
Number of slices & 126 & 106 & 50 & 92\\
Additional parameters & n/a & n/a & 1,450 ms label period;\newline  2,025 ms pulse label;\newline  24 control-label pairs & n/a\\
Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s\\
\bottomrule
\insertTableNotes

@@

The MRI scan protocol comprised of the following sequences: a
T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin
labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan,
and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging
(EPI) sequence. The DWI sequence was not used for the present study.

@@

The raw DICOM files acquired from the scanning procedure were converted
to NIfTI (Neuroimaging Informatics Technology Initiative) format using
Chris Rorden's \texttt{dcmniix} tool. SWI magnitude data files were then
used to create subject-specific brain masks that would not erode the SSS
during QSM processing, an issue faced by our group in the past (X. Li et
al. 2016). A step-by-step summary of the pipeline used is shown in
Figure~\ref{fig-sample}.

@@

\includegraphics{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}

@@

\caption{\label{fig-graph}Pipeline for generating subject-specific brain
masks that include the superior sagittal sinus (SSS). Initial steps
involved (1) creating brain a mask from the magnitude of the fifth echo
of the susceptibility weighted scan. Subsequently, the brain mask is
dilated and then (2) utilized in conjunction with a quantitative
susceptibility mapping (QSM) script to generate a preliminary QSM image.
Further refinement involved (3) segmenting the SSS from the QSM image
manually to create a tissue mask of the SSS region. Finally, (4) the
vascular mask of the SSS is integrated with the initial brain mask,
forming the comprehensive brain mask essential for obtaining
susceptibility data that includes the SSS.}

@@

\includegraphics{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}

@@

\caption{\label{fig-sample}An example of subject imaging data. A sample
sagittal, coronal, and axial slice is displayed for each image. (a) The
1st echo of the magnitude susceptibility weighted imaging sequence; (b)
the final quantitative susceptibility mapping image; and (c) the
paramagnetic component isolated from the quantitative susceptibility
map. The bars in (b) and (c) indicates the range of susceptibility
\(\chi\) values.}

@@

Once the mean susceptibility values of the SSS and CCV were obtained
from the subjects' QSM images and paramagnetic maps, venous oxygen
saturation (SvO\textsubscript{2}) was calculated with the following
equation (Berg et al. 2021):

@@

Statistical analysis of the acquired data was performed using R and
RStudio (v. 2023.09.1 Build 494) (R Core Team 2022; RStudio Team, n.d.).
Mean and standard deviation values are reported for most statistics,
unless specified otherwise. A paired Student's t-test was used to
determine statistical significance (p \textless0.05) between two
parameters (e.g.~\(\chi\) values between venous structures).

@@

\begin{longtable}[t]{lc}

\caption{\label{tbl-dem}Demographic and clinical characteristic of the
study sample.}

@@

\includegraphics{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
true
]{sn-jnl}

@@

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

@@

\raggedbottom
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{\textbf{Figure}}
\else
  \newcommand\figurename{\textbf{Figure}}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{\textbf{Table}}
\else
  \newcommand\tablename{\textbf{Table}}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\captionsetup{labelsep=period}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage[]{natbib}
\bibliographystyle{plainnat}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates},
  pdfauthor={, , , and },
  pdfkeywords={Quantitative Susceptbility
Mapping, Preterm, Newborn, Cerebral Venous Oxygen Saturation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

% author setup
\author[1,2]{\fnm{Thomas Gavin} \sur{Carmichael}}\email{tgcarmichael@outlook.com}\author[3]{\fnm{Alexander} \sur{Rauscher}}\email{rauscher@physics.ubc.ca}\author[2,3]{\fnm{Ruth E} \sur{Grunau}}\email{rgrunau@mail.ubc.ca}\author*[2,3]{\fnm{Alexander Mark} \sur{Weber}}\email{aweber@bcchr.ca}
% affil setup
\affil[1]{, \orgname{Integrated Sciences, The University of British
Columbia, Vancouver, BC, Canada}}
\affil[2]{, \orgname{BC Children's Hospital Research Institute, The
University of British Columbia, Vancouver, BC, Canada}}
\affil[3]{, \orgname{Pediatrics, The University of British Columbia,
Vancouver, BC, Canada}}

@@

\abstract{\textbf{Background}: Quantitative susceptibility mapping
(QSM), an magnetic resonance imaging (MRI) modality sensitive to
deoxyhemoglobin, is a promising method for measuring cerebral
oxygenation in human neonates. Paramagnetic sources, like
deoxyhemoglobin, however, can be obscured by diamagnetic sources such as
water and myelin. This study evaluated whether QSM images, or isolated
paramagnetic components, are more accurate for measuring oxygenation of
cerebral veins of preterm neonates, and explored oxygenation differences
between the major cerebral veins.

\textbf{Methods}: 19 preterm neonates were scanned on at term equivalent
age on a 3T MRI using a multi-echo susceptibility-weighted imaging
sequence. Susceptibility values were calculated from QSM images to
determine oxygen saturation (SvO\textsubscript{2}) in the superior
sagittal sinus (SSS) and central cerebral veins (CCV). The paramagnetic
components of QSM images were isolated, and SvO\textsubscript{2} values
were recalculated.

\textbf{Results}: The mean SvO\textsubscript{2} values from QSM were
72.4\% (SD, 3.4\%) for the SSS and and 68.7\% (SD, 3.5\%) for the CCV.
SvO\textsubscript{2} values for paramagnetic components were 58.1\% (SD,
7.3\%) for the SSS and 57.7\% (SD, 7.0\%) for the CCV.

\textbf{Conclusion}: While paramagnetic component decomposition yielded
more accurate SSS values, it increased variability. No significant
oxygenation differences were found between the SSS and CCV, contrasting
with prior studies.}

@@

\textsuperscript{1} Integrated Sciences, The University of British
Columbia, Vancouver, BC, Canada\\
\textsuperscript{2} BC Children's Hospital Research Institute, The
University of British Columbia, Vancouver, BC, Canada\\
\textsuperscript{3} Pediatrics, The University of British Columbia,
Vancouver, BC, Canada

\textsuperscript{*} Correspondence:
\href{mailto:aweber@bcchr.ca}{Alexander Mark Weber
\textless{}aweber@bcchr.ca\textgreater{}}

\textbf{Category of Study:} basic science

\subsection{Impact Statement}\label{impact-statement}

\begin{itemize}
\tightlist
\item
  This study evaluated the accuracy of using QSM and its paramagnetic
  components to measure cerebral oxygenation in neonates.
\item
  By comparing susceptibility-derived oxygen saturation
  (SvO\textsubscript{2}) in the superior sagittal sinus (SSS) and
  central cerebral veins (CCV), it adds to the field of neonatal
  cerebral oxygenation measurement.
\item
  Decomposing QSM into paramagnetic components shows potential for
  improving SvO\textsubscript{2} accuracy, particularly in the SSS,
  though variability remains a challenge.
\item
  The results suggest no significant oxygenation difference between the
  SSS and CCV, contrasting with previous findings, indicating a need for
  further research on neonatal venous oxygenation.
\end{itemize}

\subsection{Acknowledgments}\label{acknowledgments}

We wish to acknowledge the work of and thank Victoria Tapics (Research
Nurse); Vicki Goh (Research Nurse); Chacko Anil (Neonatologist); and
Michael A. Sargent (Radiologist).

\subsection{Funding}\label{funding}

Authors AMW and REG were co-primary applicants for a BC Children's
Hospital Research Institute - Brain, Behaviour and Development Catalyst
Grant (\$20,000). AMW was supported by an establishment award from
BCCHRI. Scanning was partly funded through a special award to AMW from
BCCHRI .

\subsection{Author contributions}\label{author-contributions}

TGC wrote the original draft, performed the formal analysis, and
contributed to methodology, validation, and visualization. AR helped
with writing, reviewing \& editing. REG helped with writing, reviewing
\& editing, and with initial funding acquisition. AMW was involved in
project administration, supervision, validation, visualization,
resources, methodology, formal analysis, funding acquisition, writing,
reviewing \& editing, conceptualization, data curation, and
investigation.

\subsection{Competing interests}\label{competing-interests}

The authors have no competing interests to declare.

\subsection{Consent statement}\label{consent-statement}

\newpage{}

@@

With advances in neonatal medical care, more infants born preterm are
surviving into childhood \citep{mckenzieScaffoldingParentingHealth2022}.
These children are at high risk of acquiring adverse neurodevelopmental
outcomes when compared to their term-born peers
\citep{twilhaarCognitiveOutcomesChildren2018}. Irregularities in early
cerebral oxygen levels have been identified as a potential source of
such delays, where too little oxygen provided during NICU care can
result in white matter injury, while too much oxygen can result in
reduced cortical connectivity \citep{rantakariEarlyOxygenLevels2021}. As
such, being able to precisely, accurately, and non-invasively measure
cerebral oxygenation is necessary for understanding and improving
neurodevelopmental outcomes in preterm neonates.

Unfortunately, there exist many challenges in measuring cerebral oxygen
metabolism in neonates. Cerebral metabolic rate of oxygen
(CMRO\textsubscript{2}) using oxygen-15 positron emission tomography
(PET) \citep{mintunBrainOxygenUtilization1984}, has been measured in
infants \citep{altmanCerebralBloodFlow1988}, and is considered the gold
standard. However, this method is invasive, requiring ionizing
radiation, which limits its suitability for neonates. A less invasive
option for evaluating brain hemodynamics is near-infrared spectroscopy
(NIRS), which uses the attenuation of near-infrared light
(\textasciitilde650--950 nm) as it passes through biological tissue
\citep{skovEstimationCerebralVenous1993}. Deoxygenated and oxygenated
hemoglobin absorb this light differently, allowing NIRS to estimate
changes in deoxyhemoglobin and oxyhemoglobin
\citep{wrayCharacterizationInfraredAbsorption1988} and thus provide an
estimate of cerebral venous oxygen saturation (SvO\textsubscript{2}).
While NIRS offers the advantage of being non-invasive and continuous
bedside monitoring, it is limited to regional assessments where the
probe is placed and is sensitive only to superficial brain tissue due to
the shallow penetration depth of near-infrared light
\citep{boasDiffuseOpticalImaging2004}.

For the preceding reasons, non-invasive MRI-based techniques are
actively being explored to assess regional and whole-brain blood
oxygenation. MRI based methods have been developed for adults
\citep{jainRapidMagneticResonance2011, luQuantitativeEvaluationOxygenation2008, xuNoninvasiveQuantificationWholebrain2009}
and are only beginning to be explored in neonates
\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jainCerebralOxygenMetabolism2014, jiangVesselspecificQuantificationNeonatal2019}.
These methods almost all rely on T2 relaxation to estimate
CSvO\textsubscript{2}
\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jiangVesselspecificQuantificationNeonatal2019}
with the exception of \citet{jainCerebralOxygenMetabolism2014} which
used susceptometry \citep{jainMRIEstimationGlobal2010}. One limitation
of these T2 relaxation methods, however, is the fact that
SvO\textsubscript{2} is often measured using a single imaging slice,
averaging values across several voxels, and only in the superior
sagittal sinus (SSS). In the case of
\citet{jainCerebralOxygenMetabolism2014}, they obtained regional and
whole-brain data, but with thick slices (5mm), and still only estimated
CSvO\textsubscript{2} in the SSS. An alternative MRI method using
quantitative susceptibility mapping (QSM) has been proposed, which can
measure SvO\textsubscript{2} regionally and across the whole-brain at
high resolution (\textless{} 1mm\textsuperscript{3} per voxel)
\citep{weberQuantitativeSusceptibilityMapping2021}. However, this method
left room for improvement, as it removed the SSS (averaging
CSvO\textsubscript{2} across the internal veins), and required an
arbitrary threshold value of 0.15 ppm in order to acquire realistic
results \citep{weberQuantitativeSusceptibilityMapping2021}.

In the present study, we set out to determine whether decomposing the
QSM image into its paramagnetic and diamagnetic components would allow
for a more accurate assessment of SvO\textsubscript{2} in the central
cerebral veins (CCV) of a cohort of preterm neonates. We also had a
secondary aim of preserving the SSS vessel in our QSM images and using
this data to determine whether a difference in oxygenation existed
between the SSS and the CCV.

@@

Participant data comes from a previous study *** . Participants
consisted of preterm neonates born between 25- and 31-weeks gestational
age (GA) who were admitted to the level III NICU at *** . Recruitment
took place over a span of one year, from February 2021 to January 2022,
facilitated by a dedicated research nurse. Parents of eligible infants
were approached by the research nurse prior to discharge from the NICU
to explain the study objectives and seek their consent for
participation. Infants meeting the criteria for inclusion were scanned
for the study if they had already been discharged from the NICU, were in
stable condition, and had reached a term equivalent age of 37 to 44
weeks GA. However, certain exclusion criteria were applied to ensure the
homogeneity and integrity of the study sample: infants were excluded if
there was clinical evidence of a congenital malformation or syndrome, a
TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic
infarction (\textgreater2 cm, Grade 4 intraventricular hemorrhage).

@@

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750
scanner (scanner software version DV26.0\_R03) equipped with a SREE
Medical Systems (Cleveland, OH, USA) single-channel neonatal head coil
(Table~\ref{tbl-mri}). The scans were conducted at the *** 's MRI
Research Facility. Prior to the scanning procedure, subjects were
carefully prepared by a research nurse. Swaddling and feeding were used
to ensure the comfort and cooperation of the subjects during the scan.
Importantly, no sedatives or invasive markers were utilized throughout
the procedure. Subjects were placed within a specially designed SREE
Medical Systems MRI compatible incubator, which facilitated both safety
and motion minimization. Molded foam was strategically positioned around
the head and body within the incubator to further restrict subject
movement. To protect against potential hearing damage, ear plugs were
employed during the scanning process. Additionally, a pulse oximeter was
affixed to the subject's foot to monitor arterial oxygen saturation and
heart rate throughout the scan.

@@

\caption{\label{tbl-mri}\textbf{Technical parameters for MR imaging
pulse sequences}}

@@

\toprule
 & T1w & T2w & pcASL & SWI\\
\midrule
Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated\\
Phase-encoding direction & Coronal & Sagittal & Axial & Axial\\
TR (ms) & 7.74 & 2,300 & 4,680 & 30.9\\
TE (ms) & 2.97 & 66.29 & 10.55 & 5 echoes; first echo: 5; echo spacing: 5.24\\
Flip angle & 12° & 90° & 111° & 20°\\
FOV (cm) & 20 & 20 & 24 & 25\\
Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256\\
In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977\\
Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2)\\
Number of slices & 126 & 106 & 50 & 92\\
Additional parameters & n/a & n/a & 1,450 ms label period;\newline  2,025 ms pulse label;\newline  24 control-label pairs & n/a\\
Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s\\
\bottomrule
\insertTableNotes

@@

The MRI scan protocol comprised of the following sequences: a
T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin
labeling (ASL) scan \citep{alsopRecommendedImplementationArterial2015},
a multi-echo susceptibility-weighted imaging scan
\citep{denkSusceptibilityWeightedImaging2010}, and a diffusion-weighted
imaging (DWI) spin-echo echo planar imaging (EPI) sequence. The DWI
sequence was not used for the present study.

@@

The raw DICOM files acquired from the scanning procedure were converted
to NIfTI (Neuroimaging Informatics Technology Initiative) format using
Chris Rorden's \texttt{dcmniix} tool
\citep{liFirstStepNeuroimaging2016}. SWI magnitude data files were then
used to create subject-specific brain masks that would not erode the SSS
during QSM processing, an issue faced by our group in the past *** . A
step-by-step summary of the pipeline used is shown in
Figure~\ref{fig-graph}.

@@

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}}

@@

\caption{\label{fig-graph}\textbf{Pipeline for generating
subject-specific brain masks that include the superior sagittal sinus
(SSS).} Initial steps involved (1) creating brain a mask from the
magnitude of the fifth echo of the susceptibility weighted scan.
Subsequently, the brain mask is dilated and then (2) utilized in
conjunction with a quantitative susceptibility mapping (QSM) script to
generate a preliminary QSM image. Further refinement involved (3)
segmenting the SSS from the QSM image manually to create a tissue mask
of the SSS region. Finally, (4) the vascular mask of the SSS is
integrated with the initial brain mask, forming the comprehensive brain
mask essential for obtaining susceptibility data that includes the SSS.}

@@

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}}

@@

\caption{\label{fig-sample}\textbf{An example of subject imaging data.}
A sample sagittal, coronal, and axial slice is displayed for each image.
(a) The 1st echo of the magnitude susceptibility weighted imaging
sequence; (b) the final quantitative susceptibility mapping image; and
(c) the paramagnetic component isolated from the quantitative
susceptibility map. The bars in (b) and (c) indicates the range of
susceptibility \(\chi\) values.}

@@

Once the mean susceptibility values of the SSS and CCV were obtained
from the subjects' QSM images or paramagnetic maps, venous oxygen
saturation (SvO\textsubscript{2}) was calculated with the following
equation \citep{bergInvestigatingEffectFlow2021}:

@@

Statistical analysis was performed using R and RStudio (v. 2023.09.1
Build 494)
\citep{rcoreteamLanguageEnvironmentStatistical2022, rstudioteamRStudioIntegratedDevelopment}.
Mean and standard deviation values are reported for most statistics,
unless specified otherwise. A paired Student's t-test was used to
determine statistical significance (p \textless0.05) between two
parameters (e.g.~\(\chi\) values between venous structures).

@@

\begin{longtable}[t]{lr}

\caption{\label{tbl-dem}\textbf{Demographic and clinical characteristic
of the study sample.}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}}
